---
title: 'Reproducible Research: Peer Assessment 1'
output:
  word_document: default
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
### 1. Code for reading in the dataset and/or processing the data
```{r}
options(warn = -1)
unzip("activity.zip")
data <- read.csv("activity.csv")

summary(data)
```

## What is mean total number of steps taken per day?
```{r}
dsteps <- aggregate(steps ~ date, 
                    data = data, 
                    FUN = sum, 
                    na.rm = TRUE)
head(dsteps)
```
### 2. Histogram of the total number of steps taken each day
```{r}
hist(dsteps$steps, 
     xlab = "Steps taken per day", 
     ylab = "Frequency", 
     main = "Steps taken each day", 
     breaks = 25, 
     col = "skyblue")
```

### 3. Mean and median number of steps taken each day
```{r, results= "hide"}
meansteps <- mean(dsteps$steps, 
                  round = 2,
                  na.rm = TRUE)
mediansteps <- median(dsteps$steps, 
                      round = 2,
                      na.rm = TRUE)
```
Mean of the total number of steps taken per day: `r meansteps`  
Median of the total number of steps taken per day: `r mediansteps`


## What is the average daily activity pattern?
### 4. Time series plot of the average number of steps taken
```{r}
isteps <- aggregate(steps ~ interval, 
                     data = data, 
                     FUN = mean, 
                     na.rm = TRUE)
plot(isteps$interval, 
     isteps$steps, 
     xlab = "Interval", 
     ylab = "Average number of steps",
     main= "Average number of steps taken",
     col="skyblue",
     type="l",
     lwd = 3)
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, results= "hide"}
maxsteps <- isteps[which.max(isteps$steps), ]$interval
```
On average across all the days in the dataset, interval number `r maxsteps` contains the maximum number of steps.

## Imputing missing values
### 6. Code to describe and show a strategy for imputing missing data
```{r, results= "hide"}
missing <- sum(is.na(data$steps))
```
Total number of missing values in the dataset: `r missing`

```{r, results= "hide"}
library(zoo)
data2 <- subset(data, select = c(1))
data2 <- na.aggregate(data2)
data3 <- cbind(data[, 2:3], data2[,1])
```

The missing values in the dataset were filled with the mean value for that 5-minute interval.

```{r}
data3NAs <- sum(is.na(data3))
summary(data3)
```

The new dataset is equal the original dataset but with missing data filled in.  Now this dataset has `r data3NAs` missing values.


### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
dsteps3 <- aggregate(steps ~ date, 
                    data = data3, 
                    FUN = sum, 
                    na.rm = TRUE)

hist(dsteps3$steps, 
     xlab = "Steps taken per day", 
     ylab = "Frequency", 
     main = "Steps taken each day", 
     breaks = 25, 
     col = "skyblue")
```

```{r, results= "hide"}
meansteps2 <- mean(data3$steps, 
                  round = 2,
                  na.rm = TRUE)
mediansteps2 <- median(data3$steps, 
                      round = 2,
                      na.rm = TRUE)
```
Mean of the total number of steps taken per day: `r meansteps2`  
Median of the total number of steps taken per day: `r mediansteps2`

## Are there differences in activity patterns between weekdays and weekends?

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(ggplot2)
data$date <- as.POSIXct(data$date, 
                        "%Y-%m-%d")
dayofweek <- weekdays(data$date)
data <- cbind(data, dayofweek)
```
```{r}
data$dtype <- sapply(data$date, 
                     function(x) {
    if (weekdays(x) == "sábado" | weekdays(x) =="domingo") 
    {y <- "Weekend"} else 
    {y <- "Weekday"}
    y
})
```
```{r}
data4 <- aggregate(steps ~ interval + dtype, data, mean, na.rm = TRUE)

ggplot(data4, aes(interval, steps)) + labs(x = "Interval", y = "Average steps", title = "Average number of steps taken per 5-minute interval") + geom_line() + facet_grid(~ dtype)
```

### 9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report